---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# daner

<!-- badges: start -->
<!-- badges: end -->

daner provides helpers for working with data published by [DANE](https://www.dane.gov.co) (Departamento Administrativo Nacional de Estadística). It's just a bunch of functions accumulated over time, for tasks that come up repeatedly when working with DANE's microdata and metadata:

* **DIVIPOLA codes**: standardize and join Colombia's official political-administrative division codes (departments and municipalities).
* **DDI metadata**: parse the DDI (Data Documentation Initiative) XML files DANE publishes alongside its microdata catalogs, to list a study's files and variables and their labels.
* **Microdata ingestion**: read the zipped SPSS (`.sav`) files DANE distributes, handling the quirks of how they're bundled (several files per zip, file names that don't match the zip name).

## Installation

daner is not on CRAN. Install the development version from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("edalfon/daner")
```

## Example

### DIVIPOLA codes

`fix_dept_cod()` and `fix_muni_cod()` normalize department/municipality codes to their canonical, zero-padded format:

```{r example-fix-cod}
library(daner)

fix_dept_cod(c(5, 25, "5"))
fix_muni_cod(c(5001, "5002"))
```

`join_divipola_muni_cod()` left-joins municipality names, department, coordinates and other DIVIPOLA fields onto any data frame that has a municipality-code column:

```{r example-join}
muni_df <- data.frame(cod = c("05001", "25001"))

join_divipola_muni_cod(muni_df, by = "cod")
```

### DDI metadata

`list_ddi_files()` and `get_ddi_vars()` read a DDI XML document (from a URL or a local file) and return the files/variables it describes as tidy data frames:

```{r example-ddi, eval = FALSE}
ddi_url <- "http://microdatos.dane.gov.co/index.php/catalog/ddi/419"

list_ddi_files(ddi_url)
get_ddi_vars(ddi_url)
```

### Ingesting zipped microdata

`ingest_zipped_sav()` unzips and reads the `.sav` file(s) bundled in a DANE microdata zip file:

```{r example-ingest, eval = FALSE}
ingest_zipped_sav("path/to/microdata.zip")
```
